### 2023-03-02 calculus.tex
Updates from Overleaf

    \judgment{
      \consistent{\TMV_1}{\TMV_1'} \\
      \consistent{\TMV_2}{\TMV_2'} \\
    }{
      \consistent{\TArrow{\TMV_1}{\TMV_2}}{\TArrow{\TMV_1'}{\TMV_2'}}
    }{TCArr} \\
  \end{mathpar}
  %
  \caption{Type consistency.}
  \label{fig:calculus-consistency}
\end{figure}

@@

\newcommand{\joinsTo}[3]{\ensuremath{\TJoin{#1}{#2} & = & #3}}
\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\TJoin{\TMV_1}{\TMV_2}}} is a \emph{partial} metafunction defined as follows:
  %
  \[\begin{array}{rcl}
    \joinsTo{\TUnknown}{\TMV}{\TMV} \\
    \joinsTo{\TMV}{\TUnknown}{\TMV} \\
    \joinsTo{\TNum}{\TNum}{\TNum} \\
    \joinsTo{\TBool}{\TBool}{\TBool} \\
    \joinsTo{(\TArrow{\TMV_1}{\TMV_2})}{(\TArrow{\TMV_1'}{\TMV_2'})}{\TArrow{(\TJoin{\TMV_1}{\TMV_1'})}{(\TJoin{\TMV_2}{\TMV_2'})}} \\
  \end{array}\]
  %
  \caption{Type join.}
  \label{fig:calculus-type-join}
\end{figure}
% TODO intuition for type join?

@@

\subsubsection{Additional Theorems}
To conclude this section, we present two more metatheorems that help ensure correctness of the
system. Though \cref{thm:calculus-marking-well-formedness} guarantees that marking does not change
the syntactic structure of a program, it makes no statement about the presence of error marks in the
resulting marked program. Following the logic of constructing marking rules, we have
\cref{thm:calculus-marking-well-ill-typed}.

@@

Finally, no less importantly, marking is deterministic. This is given by
\cref{thm:calculus-marking-unicity}.
%
\begin{theorem}[name=Marking Unicity] \
  \label{thm:calculus-marking-unicity}
  \begin{enumerate}
    \item If $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV_1}{\TMV_1}$ and
      $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV_2}{\TMV_2}$, then $\ECMV_1 = \ECMV_2$ and $\TMV_1 =
      \TMV_2$.
    \item If $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV_1}{\TMV}$ and
      $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV_2}{\TMV}$, then $\ECMV_1 = \ECMV_2$.
  \end{enumerate}
\end{theorem}
%
Together, totality and unicity give that marking may be considered and implemented as a total
function. Indeed, given the algorithmic nature of bidirectional typing, it is fairly easy to extract
an implementation from the rules.

@@

First, we extend the syntax of types and the accompanying notions of consistency and join, which
closely mirror the definitions for arrow types. Additionally, analogous to the notion of matched
arrow types, the judgment $\matchedProd{\TMV}{\TMV_1}{\TMV_2}$ establishes that $\TMV$ may be
considered the binary product type $\TProd{\TMV_1}{\TMV_2}$ [cite?].

@@

We now consider how the calculus might be extended to richer typing features such as parametric
polymorphism. Specifically, we consider System F-style explicit polymorphism.

\subsection{Mechanization}
\label{sec:calculus-agda}

>>>>

    \judgment{
      \consistent{\TMV_1}{\TMV_1'} \\
      \consistent{\TMV_2}{\TMV_2'} \\
    }{
      \consistent{\TArrow{\TMV_1}{\TMV_2}}{\TArrow{\TMV_1'}{\TMV_2'}}
    }{TCArr}
  \end{mathpar}
  \vspace{-10px}
  \caption{Type consistency.}
  \label{fig:calculus-consistency}
\end{figure}

@@

\newcommand{\joinsTo}[3]{\ensuremath{\TJoin{#1}{#2} & = & #3}}
\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\TJoin{\TMV_1}{\TMV_2}}} is a \emph{partial} metafunction defined as follows:
  %
  \[\begin{array}{rcl}
    \joinsTo{\TUnknown}{\TMV}{\TMV} \\
    \joinsTo{\TMV}{\TUnknown}{\TMV} \\
    \joinsTo{\TNum}{\TNum}{\TNum} \\
    \joinsTo{\TBool}{\TBool}{\TBool} \\
    \joinsTo{(\TArrow{\TMV_1}{\TMV_2})}{(\TArrow{\TMV_1'}{\TMV_2'})}{\TArrow{(\TJoin{\TMV_1}{\TMV_1'})}{(\TJoin{\TMV_2}{\TMV_2'})}}
  \end{array}\]
  \vspace{-10px}
  \caption{Type join.}
  \label{fig:calculus-type-join}
\end{figure}
% TODO intuition for type join?

@@

\subsubsection{Additional Metatheorems}
To conclude this section, we present two more metatheorems that help ensure correctness of the
system. Though \cref{thm:calculus-marking-well-formedness} guarantees that marking does not change
the syntactic structure of a program, it makes no statement about the presence of error marks in the
resulting marked program. Following the logic of constructing marking rules, we have
\cref{thm:calculus-marking-well-ill-typed}.

@@

Finally, no less importantly, marking is deterministic. This is given by
\cref{thm:calculus-marking-unicity}.
%
\begin{theorem}[name=Marking Unicity] \
  \label{thm:calculus-marking-unicity}
  \begin{enumerate}
    \item If $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV_1}{\TMV_1}$ and
      $\ctxSynFixedInto{\ctx}{\EMV}{\ECMV_2}{\TMV_2}$, then $\ECMV_1 = \ECMV_2$ and $\TMV_1 =
      \TMV_2$.
    \item If $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV_1}{\TMV}$ and
      $\ctxAnaFixedInto{\ctx}{\EMV}{\ECMV_2}{\TMV}$, then $\ECMV_1 = \ECMV_2$.
  \end{enumerate}
\end{theorem}
%
Together, totality and unicity give that marking may be considered and implemented as a total
function. Indeed, given the algorithmic nature of bidirectional typing, it is fairly direct to extract
an implementation from these rules.

@@

First, we extend the syntax of types and the accompanying notions of consistency and join, which
closely mirror the definitions for arrow types. Additionally, analogous to the notion of matched
arrow types, the judgment $\matchedProd{\TMV}{\TMV_1}{\TMV_2}$ establishes that $\TMV$ may be
considered the binary product type $\TProd{\TMV_1}{\TMV_2}$ [cite?].\todo{cite}

@@

We now consider how the calculus might be extended to richer typing features such as parametric
polymorphism. Specifically, we consider System F-style explicit polymorphism.\todo{write me}

\subsection{Agda Mechanization}
\label{sec:calculus-agda}
